---
title: "Untitled"
output: html_document
---




```{r setup, include=FALSE}
library('knitr')
knitr::opts_chunk$set(cache=TRUE)
```


Let's simulate data with one covariance matrix, and $\omega$ fixed at 2.
We can have a look at the them first. Here, for the inference we will still scale by all the $\omegas$ and later, we will compare with the estimates that come from using the true prior in both MASH and MASHNobaseline.

```{r}
rm(list=ls())
library('mashr')
set.seed(123)
c=chat_sim_fsingle_fixedomega(n = 1000,d = 8,omega = 2,esd = 1,mu = FALSE)
saveRDS(c,"chatmufixedomega.rds")
t=c$t;b=c$chat;se=c$shat
length(which(rowSums(c$beta)!=0))

###Here we're just going to use the 'real associations'
index=which(rowSums(c$beta)!=0)
strongt=t[index,]
R=ncol(strongt)
v.j=se
```

Here we initialize with empirical covariance and run MASH:

```{r}
library("ExtremeDeconvolution")
init.cov.list=list()
init.cov.list[[1]]=cov(strongt)
#head(init.cov.list)
mean.mat=matrix(rep(0,R*R),ncol=R,nrow=R)  
ydata= strongt##train on the max like estimates
xamp=1
xcovar=init.cov.list
fixmean=TRUE     
ycovar=v.j     
xmean=mean.mat   
projection=list();for(l in 1:nrow(strongt)){projection[[l]]=diag(1,R)}
e=extreme_deconvolution(ydata=ydata,ycovar=ycovar,xamp=xamp,xmean=xmean,xcovar=init.cov.list,fixmean=T,projection=projection)
se.train=se


covlist=e$xcovar

train.b=data.frame(strongt);se.train = se;covmat = covlist;pen = 1;L = diag(1,R);J=nrow(train.b)
lik.mat=t(sapply(seq(1:J),function(j){
  V.gp.hat=diag(se.train[j,])^2;
  b.mle=train.b[j,]
  dmvnorm(x=b.mle, sigma=(covlist[[1]]+V.gp.hat),log=TRUE)}))

dim(lik.mat)
sum(log(exp(lik.mat)))
j=10
lik.mat[10]
V.gp.hat=diag(se.train[j,])^2;
dmvnorm(x=train.b[j,], sigma=(covlist[[1]]+V.gp.hat),log=TRUE)
mean(lik.mat)
e$avgloglikedata
```



Here we initialize with the Identity and run MASH:

```{r}
rm(init.cov.list)
rm(covlist)
init.cov.list=list()
init.cov.list[[1]]=diag(1,R)
#head(init.cov.list)
mean.mat=matrix(rep(0,R*R),ncol=R,nrow=R)  
ydata= strongt##train on the max like estimates
xamp=1
xcovar=init.cov.list
fixmean=TRUE     
ycovar=v.j     
xmean=mean.mat   
projection=list();for(l in 1:nrow(strongt)){projection[[l]]=diag(1,R)}
e=extreme_deconvolution(ydata=ydata,ycovar=ycovar,xamp=xamp,xmean=xmean,xcovar=init.cov.list,fixmean=T,projection=projection)
se.train=se

covlist=e$xcovar

train.b=data.frame(strongt);se.train = se;covmat = covlist;pen = 1;L = diag(1,R);J=nrow(train.b)
lik.mat=t(sapply(seq(1:J),function(j){
  V.gp.hat=diag(se.train[j,])^2;
  b.mle=train.b[j,]
  dmvnorm(x=b.mle, sigma=(covlist[[1]]+V.gp.hat),log=TRUE)}))

dim(lik.mat)
sum(log(exp(lik.mat)))
j=10
lik.mat[10]
V.gp.hat=diag(se.train[j,])^2;
dmvnorm(x=train.b[j,], sigma=(covlist[[1]]+V.gp.hat),log=TRUE)
mean(lik.mat)
e$avgloglikedata

```

Here we initialize with the Truth and run MASH:

```{r}
rm(init.cov.list)
rm(covlist)
init.cov.list=list();init.cov.list[[1]]=c$covmat
#head(init.cov.list)
mean.mat=matrix(rep(0,R*R),ncol=R,nrow=R)  
ydata= strongt##train on the max like estimates
xamp=1
xcovar=init.cov.list
fixmean=TRUE     
ycovar=v.j     
xmean=mean.mat   
projection=list();for(l in 1:nrow(strongt)){projection[[l]]=diag(1,R)}
e=extreme_deconvolution(ydata=ydata,ycovar=ycovar,xamp=xamp,xmean=xmean,xcovar=init.cov.list,fixmean=T,projection=projection)
se.train=se

covlist=e$xcovar

train.b=data.frame(strongt);se.train = se;covmat = covlist;pen = 1;L = diag(1,R);J=nrow(train.b)
lik.mat=t(sapply(seq(1:J),function(j){
  V.gp.hat=diag(se.train[j,])^2;
  b.mle=train.b[j,]
  dmvnorm(x=b.mle, sigma=(covlist[[1]]+V.gp.hat),log=TRUE)}))

dim(lik.mat)
sum(log(exp(lik.mat)))
j=10
lik.mat[10]
V.gp.hat=diag(se.train[j,])^2;
dmvnorm(x=train.b[j,], sigma=(covlist[[1]]+V.gp.hat),log=TRUE)

mean(lik.mat)
e$avgloglikedata
```

And here we run with the truth:
```{r}
rm(list=ls())
c=readRDS("chatmufixedomega.rds")
t=c$t;b=c$chat;se=c$shat
length(which(rowSums(c$beta)!=0))

###Here we're just going to use the 'real associations'
index=which(rowSums(c$beta)!=0)
strongt=t[index,]
R=ncol(strongt)
v.j=se
```

Here we just use the truth:

```{r}
covlist=list();covlist[[1]]=c$covmat
train.b=data.frame(strongt);se.train = se;covmat = covlist;pen = 1;L = diag(1,R);J=nrow(train.b)
lik.mat=t(sapply(seq(1:J),function(j){
  V.gp.hat=diag(se.train[j,])^2;
  b.mle=train.b[j,]
  dmvnorm(x=b.mle, sigma=(covlist[[1]]+V.gp.hat),log=TRUE)}))

dim(lik.mat)
sum(log(exp(lik.mat)))
j=10
lik.mat[10]
V.gp.hat=diag(se.train[j,])^2;
dmvnorm(x=train.b[j,], sigma=(covlist[[1]]+V.gp.hat),log=TRUE)
```

