---
title: "johnsdatamashmedian"
output: html_document
---
```{r}
rm(list=ls())

library('mash')


t=read.table("tbatchcorrected.txt")
b=read.table("betabatchcorrected.txt")
se=read.table("sebetabatchcorrected.txt")
#Now wfull will be the median
wfull=t(apply(b,1,function(x){x-median(x)}))
##show that they sum to 0
```

Now we proceed as in mash, selecting the strongest deviations from the median:
```{r}

s.j=se/se
absmat=abs(t(apply(t,1,function(x){x-median(x)})))
tmedian=(t(apply(t,1,function(x){x-median(x)})))

index=which(rowSums(absmat>4)>0)
length(index)


sparsestrongt=tmedian[index,]
strongprojectedt=sparsestrongt
```



```{r filewrite,eval=T}

write.table(sparsestrongt,"sparsestrongt.txt",col.names = FALSE,row.names=FALSE)

```


Now, we need to project into the centered space to estimate the covariance matrix of the true deviations, using the full L since $v$ will be R, and not $R-1$.


```{r,eval=T}
system('/Users/sarahurbut/miniconda3/bin/sfa -gen sparsestrongt.txt -g 137 -k 5 -n 8 i -o sparseFmediancentered')
A="sparseFmediancentered"

factor.mat=as.matrix(read.table("sparseFmediancentered_F.out"))
lambda.mat=as.matrix(read.table("sparseFmediancentered_lambda.out"))

library("ExtremeDeconvolution")

max.step=deconvolution.em.with.bovy(t.stat = strongprojectedt,factor.mat = factor.mat,v.j = s.j,lambda.mat = lambda.mat,K = 3,P = 3 )

##Show how the ranks is preserved
apply(max.step$true.covs,1,function(x){qr(x)$rank})


A="sparseoldmash"

covmash=compute.hm.mash.covmat.all.max.step.fixed.omega(omega = 2,Q = 5,lambda.mat = lambda.mat,factor.mat = factor.mat,max.step = max.step,zero = TRUE,power = 2,A,bma = TRUE)$cov

sapply(covmash,max)

##Show how the ranks is preserved
apply(max.step$true.covs,1,function(x){qr(x)$rank})


all.equal(covmash[[2]]/max(diag(covmash[[2]])),max.step$true.covs[1,,]/max(diag(max.step$true.covs[1,,])))
all.equal(covmash[[3]]/max(diag(covmash[[3]])),max.step$true.covs[3,,]/max(diag(max.step$true.covs[3,,])))
all.equal(covmash[[9]]/max(diag(covmash[[9]])),max.step$true.covs[2,,]/max(diag(max.step$true.covs[2,,])))

length(covmash)
# 
# 
covmash=compute.hm.covmat.all.max.step(max.step = max.step,b.hat = wfull,se.hat = se,t.stat = strongprojectedt,Q = 5,lambda.mat = lambda.mat,A = "sparseoldmash",factor.mat = factor.mat,zero = T,power = 2)$covmat

all.equal(covmash[[2]]/max(diag(covmash[[2]])),max.step$true.covs[1,,]/max(diag(max.step$true.covs[1,,])))
all.equal(covmash[[3]]/max(diag(covmash[[3]])),max.step$true.covs[3,,]/max(diag(max.step$true.covs[3,,])))
all.equal(covmash[[9]]/max(diag(covmash[[9]])),max.step$true.covs[2,,]/max(diag(max.step$true.covs[2,,])))

```

Now, proceed as always, estimating likelihood and computing posteriors.

```{r mashmethod, eval=F}
A="sparseoldmash"
compute.hm.train.log.lik.pen(train.b = wfull,se.train = se,covmat = covmash,A=A,pen=1)


pis=readRDS(paste0("pis",A,".rds"))$pihat

weightedquants=lapply(seq(1:nrow(wfull)),function(j){total.quant.per.snp(j,covmash,b.gp.hat=wfull,se.gp.hat = se,pis,A=A,checkpoint = FALSE)})
```



Let's look at the 'learned' covariance matrix of each as well as the true covariance matrices of each, and see how they comapre to the true configurations. Recall that the true configurations looked like this:


```{r}
pis=readRDS("pissparseoldmash.rds")$pihat
pi.mat.old.mash=matrix(pis[-length(pis)],ncol=18,byrow=TRUE)
covmash=readRDS("covmatsparseoldmash.rds")

library('gplots')
library('colorRamps')

for(i in 2:9){
  x=covmash[[i]]
  heatmap.2(x/max(diag(x)),Rowv = NULL,Colv = NULL,dendrogram = "none",col=blue2green,density.info = "none",trace="none",main=paste0("CovMash",round(colSums(pi.mat.old.mash)[i],2)))}



```


Test to show our calculations are accurate:

```{r,eval=T,echo=T}
mash.means=as.matrix(read.table("sparseoldmashposterior.means.txt")[,-1])
lfsr.mash=as.matrix(read.table("sparseoldmashlfsr.txt"))[,-1]
cov=readRDS("covmatsparseoldmash.rds")
pis=readRDS("pissparseoldmash.rds")$pihat
pi.mat.old.mash=matrix(pis[-length(pis)],ncol=18,byrow=TRUE)
#pisnobase=readRDS("pissparseLwithED.rds")$pihat
#pi.mat.nobase=matrix(pisnobase[-length(pisnobase)],ncol=18,byrow=TRUE)
barplot(colSums(pi.mat.old.mash))
#barplot(colSums(pi.mat.nobase))

w=data.frame(wfull)
R=ncol(w)
j=10

arrays=post.array.per.snp.no.baseline(j = 10,covmat = cov,b.gp.hat = w,se.gp.hat = se,L = diag(1,R))
mash.arrays=post.array.per.snp(j = 10,covmat = cov,b.gp.hat = w,se.gp.hat = se)
##show same algebra


k=5

V.gp.hat=diag(se[j,])^2

b.mle = as.vector(w[j,])

lapply(seq(1:5),function(x){all.equal(arrays[[x]],mash.arrays[[x]])})
L=diag(1,R)
LSigL=L%*%cov[[k]]%*%t(L)
LVL=L%*%V.gp.hat%*%t(L)

LSigL_list=lapply(cov,function(x){L%*%x%*%t(L)})
identical(LSigL_list[[k]],LSigL)
identical(as.numeric(arrays$post.means[k,]),as.numeric(post.mean.with.proj(b.mle = t(b.mle),tinv = solve(LVL+L%*%cov[[k]]%*%t(L)),U.k = cov[[k]],L = L)))
identical(as.numeric(arrays$post.covs[k,]),as.numeric(diag(post.cov.with.proj(tinv = solve(LVL+L%*%cov[[k]]%*%t(L)),U.k = cov[[k]],L = L))))




b.mle=as.vector(t(w[j,]))##turn i into a R x 1 vector
V.gp.hat=diag(se[j,])^2
V.gp.hat.inv <- solve(V.gp.hat)
U.gp1kl= (post.b.gpkl.cov(V.gp.hat.inv, cov[[k]]))
  
identical(as.numeric(mash.arrays$post.means[k,]),as.numeric(post.b.gpkl.mean(b.mle = b.mle,V.gp.hat.inv = V.gp.hat.inv,U.gp1kl = U.gp1kl)))

identical(as.numeric(mash.arrays$post.covs[k,]),as.numeric(diag(U.gp1kl)))



log.lik.snp=log.lik.func(b.mle,LVL,LSigL_list)
log.lik.minus.max=log.lik.snp-max(log.lik.snp)
exp.vec=exp(log.lik.minus.max)
post.weights=t(exp.vec*pis/sum(exp.vec*pis))

all.equal(as.numeric(post.weights%*%arrays$post.means),as.numeric(mash.means[j,]))

```


Let's also plot pairwise sharing by magnitude and sign:



```{r pairwisesharing}
library("mashr")
post.means=as.matrix(read.table("sparseoldmashposterior.means.txt")[,-1])
lfsr.mash=as.matrix(read.table("sparseoldmashlfsr.txt"))[,-1]

se.matched=as.matrix(read.table("sebetabatchcorrected.txt"))
strain.names=read.table("StrainNames.txt",header = T,stringsAsFactors = F)[,1]
pm.mash.beta=post.means*se.matched
colnames(pm.mash.beta)=colnames(lfsr.mash)=strain.names
lfsr.mash.sig=lfsr.mash[rowSums(lfsr.mash<0.05)>0,]##only 137,223 are significant in at least one subgroup
pm.mash.sig=pm.mash.beta[rowSums(lfsr.mash<0.05)>0,]

signheatmap=compute.sharing.by.sign(lfsr.mash = lfsr.mash.sig,thresh = 0.05,pm.mash.beta = pm.mash.sig)
signheatmap[lower.tri(signheatmap)] <- NA
magheatmap=compute.mag.by.sharing(lfsr.mash = lfsr.mash.sig,thresh = 0.05,pm.mash.beta = pm.mash.sig)
magheatmap[lower.tri(magheatmap)] <- NA


library('colorRamps')
library('corrplot')
library(gplots)
library(ggplot2)

class(signheatmap)



library('lattice')

clrs <- colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(64)
#clrs[63:64] <- "darkviolet"
lat=signheatmap
lat[lower.tri(lat)] <- NA
print(levelplot(lat,col.regions = clrs,xlab = "",ylab = "",colorkey = TRUE,main="PairwiseSharingBySign"))


heatmap.2(magheatmap,Rowv=FALSE,Colv=FALSE,
          symm=TRUE,dendrogram="none",density="none",trace="none",#col=redblue,
          col=blue2green,main=paste0("Pairwise Sharing by Magnitude"),
          cexRow=0.6,cexCol=2,cex.main=0.5,#breaks=seq(0.35,1,0.01),
          labCol=NA)

library('lattice')

clrs <- colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(64)
#clrs[63:64] <- "darkviolet"
absmagheatmap=compute.mag.by.sharing(lfsr.mash = lfsr.mash.sig,thresh = 0.05,pm.mash.beta = abs(pm.mash.sig))
absmagheatmap[lower.tri(absmagheatmap)] <- NA
lat=absmagheatmap
lat[lower.tri(lat)] <- NA
print(levelplot(lat,col.regions = clrs,
                xlab = "",ylab = "",colorkey = TRUE,main="PairwiseSharingByMagnitude"))
```

Now, let's look at sharing my asbolute value of magntiude:
```{r}

heatmap.2(signheatmap,Rowv=FALSE,Colv=FALSE,
          symm=TRUE,dendrogram="none",density="none",trace="none",#col=redblue,
          col=blue2green,main=paste0("Pairwise Sharing by Sign"),
          cexRow=0.6,cexCol=0.5,cex.main=0.5)
#,breaks=seq(0.7,1,0.01))

heatmap.2(1-signheatmap,Rowv=FALSE,Colv=FALSE,
          symm=TRUE,dendrogram="none",density="none",trace="none",#col=redblue,
          col=blue2green,main=paste0("Pairwise Sharing by Opposite Sign"),
          cexRow=0.6,cexCol=0.5,cex.main=0.5)
#,breaks=seq(0.7,1,0.01))



absmagheatmap=compute.mag.by.sharing(lfsr.mash = lfsr.mash.sig,thresh = 0.05,pm.mash.beta = abs(pm.mash.sig))
absmagheatmap[lower.tri(absmagheatmap)] <- NA
heatmap.2(absmagheatmap,Rowv=FALSE,Colv=FALSE,
          symm=TRUE,dendrogram="none",density="none",trace="none",#col=redblue,
          col=blue2green,main=paste0("Pairwise Sharing by abs(Magnitude)"),
          cexRow=0.6,cexCol=2,cex.main=0.5,#breaks=seq(0.35,1,0.01),
          labCol=NA)

heatmap.2(magheatmap,Rowv=FALSE,Colv=FALSE,
          symm=TRUE,dendrogram="none",density="none",trace="none",#col=redblue,
          col=blue2green,main=paste0("Pairwise Sharing by Magnitude"),
          cexRow=0.6,cexCol=2,cex.main=0.5,#breaks=seq(0.35,1,0.01),
          labCol=NA)
```





